---
title: "Generalized Nadaraya-Watson (2d) Kernel Regression"
author: "Jordan Schupbach"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Generalized Nadaraya-Watson (2d) Kernel Regression}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- {{{ Setup -->
```{r setup}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
<!-- }}} Setup -->

## Introduction
This vignette demonstrates how to use the `mixedcurve` package to fit a
generalized Nadaraya-Watson kernel regression model to two-dimensional Poisson
data.

## Example Usage
Let's start by simulating some functional Poisson data with group effects to
fit a generalized Nadaraya-Watson kernel regression model to. We will make some
modifications to the Cuevas et al. m3 curve for this purpose.

<!-- {{{ Simulate data -->

```{r, fig.filename="m3_2d_gdata.png", fig.width=10, fig.height=10, out.width="98%", message=FALSE, warning=FALSE}

set.seed(123)
# 1. Define the true curves
tf <- function(t, i) {
  # Define the rate function for Poisson data at time t
  20 * mixedcurve::m2(t, i)
}
n <- 7000
set.seed(123)
df1 <- mixedcurve::gen_hfanova_data(
  f = tf,
  bounds = list(c(0, 1),
                c(0, 1)),
  n = c(n, 2),
  sigmas = c(0.1, 0),
  ndim = 2,
  ngrp = 1,
  px = NULL,
  pxargs = NULL,
  family = "poisson",
  white_noise = TRUE
)
color_values <- cut(df1$y, breaks = 100, labels = FALSE)
color_scale <- viridis::viridis(100)
# png("./gnw_2d_poisson_data.png", width = 800, height = 800, res = 100)
mixedcurve::dark_mode()
plot(df1$x1, df1$x2,
  col = color_scale[color_values],
  pch = 20,
  xlab = "x1", ylab = "x2",
  main = "Modified 2D M3 Poisson Data"
)


```

<!-- }}} Simulate data -->

Now, we can fit the generalized Nadaraya-Watson kernel regression model using
the `lpk` function from the `mixedcurve` package. We will specify the
bandwidth, kernel type, degree, and use the formula `y ~ K_h(x * y)` to
indicate that we want to fit a generalized local polynomial kernel model across
domain `x * y`, i.e. the two-dimensional plane.

<!-- {{{ Fit GNW model -->

```{r, fig.filename="m3_2d_gfit.png", fig.width=20, fig.height=10, out.width="98%", message=FALSE, warning=FALSE}

# 4. Fit GNW kernel regression model (in parallel)
bandwidth <- 0.03
nq <- 12
qseq <- as.matrix(expand.grid(
  seq(0.0, 1.0, length.out = nq),
  seq(0.0, 1.0, length.out = nq)
))
cl <- parallel::makeCluster(parallel::detectCores() - 1)
invisible(parallel::clusterEvalQ(cl, library(mixedcurve)))
parallel::clusterExport(cl, varlist = c("df1", "bandwidth", "qseq"))
glpk1 <- mixedcurve::glpk(y ~ K_h(x1 * x2),
  queries = qseq,
  data = df1,
  degree = 0,
  kernel = mixedcurve::gauss_kern,
  family = "poisson",
  h = bandwidth,
  kthresh = 0.1,
  parallel = TRUE,
  cl = cl
)
parallel::stopCluster(cl)
fits <- unlist(lapply(glpk1[[1]], function(elmt) { exp(elmt$coefs) }))
all_values <- c(fits, exp(tf(expand.grid(
  seq(0.0, 1.0, length.out = nq),
  seq(0.0, 1.0, length.out = nq)
), 1)))
breaks <- seq(min(all_values), max(all_values), length.out = 101)
colors <- viridis::viridis(100)
par(mfrow = c(1, 2))
mixedcurve::dark_mode()
image(matrix(fits, nq, nq, byrow = TRUE),
  col = colors, breaks = breaks,
  main = "GNW fit"
)
image(exp(matrix(tf(expand.grid(
  seq(0.0, 1.0, length.out = nq),
  seq(0.0, 1.0, length.out = nq)
), 1), nq, nq, byrow = TRUE)),
  col = colors, breaks = breaks,
  main = "True curve"
)

#
```

<!-- }}} Fit GNW model -->

## WY Adjusted Test

We can also compute Westfall-Young adjusted p-values for the
fitted model using

<!-- {{{ WY adjusted p-values  -->

```{r, fig.filename="doppler_wy.png", fig.width=20, fig.height=10, out.width="98%", message=FALSE, warning=FALSE}

cl <- parallel::makeCluster(parallel::detectCores() - 1)
invisible(parallel::clusterEvalQ(cl, library(mixedcurve)))
parallel::clusterExport(cl, varlist = c("df1", "bandwidth", "qseq"))
wy_pvals <- mixedcurve::wy_full(
  dataf = df1, xseq = qseq, nperm = 50,
  gen_pvals_fun = function(data, xseq) {
    lpk_res <- mixedcurve::lpk(
      y ~ K_h(x1 * x2), queries = xseq, data = data, degree = 0,
      kernel = mixedcurve::box_kern,
      h = bandwidth, parallel = FALSE,
      alternative_hypothesis = y ~ K_h(x1 * x2 | -1),
      kthresh = 0.1
    )
    as.numeric(do.call(cbind, lapply(lpk_res[[1]], function(elmt) {
      elmt$pvals
    })))
  },
  gen_perm_fun = function(data) {
    data$y <- sample(data$y, replace = FALSE)
    data
  },
  cl = cl
)
parallel::stopCluster(cl)
color_values <- cut(wy_pvals, breaks = 100, labels = FALSE)
colors <- viridis::viridis(100)
mixedcurve::dark_mode()
par(mfrow = c(1, 2))
image(matrix(fits, nq, nq, byrow = TRUE),
  col = colors, breaks = breaks,
  axes = FALSE, main = paste("Nadaraya-Watson Fit Cuevas M2 (Group 1)")
)
image(matrix(wy_pvals, nrow = nq, ncol = nq, byrow = TRUE),
  col = colors,
  axes = FALSE,
  main = "Westfall-Young Adjusted P-values"
)
```

<!-- }}} WY adjusted p-values  -->

