---
title: "Nadaraya-Watson (2d) Kernel Regression w/ Covariate"
author: "Jordan Schupbach"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Nadaraya-Watson (2d) Kernel Regression w/ Covariate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- {{{ Setup -->
```{r setup}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
<!-- }}} Setup -->

## Introduction
This vignette demonstrates how to use the `mixedcurve` package to fit a
Nadaraya-Watson kernel regression model to one-dimensional data with a single covariate.

## Example Usage
Let's start by simulating some data according to the m2 Cuevas et al. function
modified to two dimensions:

<!-- {{{ Generate data -->

```{r, fig.filename="m3_2d_w_cov_gdata.png", fig.width=30, fig.height=10, out.width="98%", message=FALSE, warning=FALSE}

set.seed(123)
# 1. Define the true curves
n <- 1000
set.seed(123)
df1 <- mixedcurve::gen_hfanova_data(
  f = mixedcurve::m2,
  bounds = list(c(0, 1),
                c(0, 1)),
  n = c(n, 1),
  sigmas = c(0.01, 0),
  ndim = 2,
  ngrp = 3,
  px = NULL,
  pxargs = NULL,
  family = "gaussian",
  white_noise = TRUE
)
color_values <- cut(df1$y, breaks = 100, labels = FALSE)
color_scale <- viridis::viridis(100)
mixedcurve::dark_mode()
par(mfrow = c(1, 3))
plot(df1$x1[df1$cov == 1], df1$x2[df1$cov == 1],
  col = color_scale[color_values[df1$cov == 1]],
  pch = 20,
  cex = 3.0,
  main = "M3 Data (Group 1)"
)
plot(df1$x1[df1$cov == 2], df1$x2[df1$cov == 2],
  col = color_scale[color_values[df1$cov == 2]],
  pch = 20,
  cex = 3.0,
  main = "M3 Data (Group 2)"
)
plot(df1$x1[df1$cov == 3], df1$x2[df1$cov == 3],
  col = color_scale[color_values[df1$cov == 3]],
  pch = 20,
  cex = 3.0,
  main = "M3 Data (Group 3)"
)
```
<!-- }}} Generate data -->

To fit the Nadaraya-Watson kernel regression model, we can use the `lpk`
function from the `mixedcurve` package. We will specify the bandwidth, kernel
type, and degree and the function `y ~ K_h(x1 * x2 | grp)` to indicate a
two-dimensional kernel regression on the 2d (product) space of `x1 * x2` with a
covariate `grp`, fitting surfaces for each group.

<!-- {{{ Fit NW model -->

```{r, fig.filename="lpk_2d_m2_w_cov_fit.png", fig.width=30, fig.height=20, out.width="98%", message=FALSE, warning=FALSE}

# Fit Nadaraya-Watson kernel regression model (in parallel)
nq <- 20
qseq <- as.matrix(expand.grid(
  seq(0.0, 1.0, length.out = nq),
  seq(0.0, 1.0, length.out = nq))
)
bandwidth <- 0.02
cl <- parallel::makeCluster(parallel::detectCores() - 1)
invisible(parallel::clusterEvalQ(cl, library(mixedcurve)))
parallel::clusterExport(cl, varlist = c("df1", "bandwidth", "qseq"))
lpk2 <- mixedcurve::lpk(y ~ K_h(x1 * x2 | cov),
  queries = qseq, data = df1, degree = 0,
  kernel = mixedcurve::box_kern, h = bandwidth, parallel = TRUE,
  kthresh = 0.1,
  cl = cl
)
parallel::stopCluster(cl)
coefs <- matrix(unlist(lapply(lpk2[[1]], function (fit) fit$coefs)), ncol = 3, byrow = TRUE)
fits <- cbind(coefs[, 1], coefs[, 1] + coefs[, 2], coefs[, 1] + coefs[, 3])
all_values <- c(fits, mixedcurve::m2(qseq, 1), mixedcurve::m2(qseq, 2), mixedcurve::m2(qseq, 3))
breaks <- seq(min(all_values), max(all_values), length.out = 101)
colors <- viridis::viridis(100)
mixedcurve::dark_mode()
par(mfrow = c(2, 3))
for (i in 1:3) {
  image(matrix(mixedcurve::m2(qseq, i), 20, 20),
    col = colors, breaks = breaks,
    axes = FALSE, main = paste("Cuevas M2 True Function (Group", i, ")")
  )
}
for (i in 1:3) {
  image(matrix(fits[, i], 20, 20),
    col = colors, breaks = breaks,
    axes = FALSE, main = paste("Nadaraya-Watson Fits Cuevas M2 Group", i)
  )
}
```

<!-- }}} Fit NW model -->

## WY Adjusted Test

We can also compute Westfall-Young adjusted p-values for the fitted model using

<!-- {{{ WY adjusted p-values  -->

```{r, fig.filename="wy_2d_glpk.png", fig.width=20, fig.height=20, out.width="98%", message=FALSE, warning=FALSE}

cl <- parallel::makeCluster(parallel::detectCores() - 1)
invisible(parallel::clusterEvalQ(cl, library(mixedcurve)))
parallel::clusterExport(cl, varlist = c("df1", "bandwidth", "qseq"))
wy_pvals <- mixedcurve::wy_full(
  dataf = df1, xseq = qseq, nperm = 50,
  gen_pvals_fun = function(data, xseq) {
    lpk_res <- mixedcurve::lpk(
      formula = y ~ K_h(x1 * x2 | cov),
      alternative_hypothesis = y ~ K_h(x1 * x2),
      queries = xseq, data = data, degree = 0,
      kernel = mixedcurve::box_kern,
      h = bandwidth, parallel = FALSE,
      kthresh = 0.1
    )
    as.numeric(do.call(cbind, lapply(lpk_res[[1]], function(elmt) {
      elmt$pvals
    })))
  },
  gen_perm_fun = function(data) {
    data$y <- sample(data$y, replace = FALSE)
    data
  },
  cl = cl
)
parallel::stopCluster(cl)
all_values <- c(coefs[,1], coefs[,1] + coefs[,2], coefs[,1] + coefs[,3])
breaks <- seq(min(all_values), max(all_values), length.out = 101)
colors <- viridis::viridis(100)
mixedcurve::dark_mode()
par(mfrow = c(2, 2))
image(matrix(coefs[,1], nq, nq, byrow = TRUE),
  col = colors, breaks = breaks,
  axes = FALSE, main = paste("LPK Fit Cuevas M2 (Group 1)")
)
image(matrix(coefs[,1] + coefs[,2], nq, nq, byrow = TRUE),
  col = colors, breaks = breaks,
  axes = FALSE, main = paste("LPK Fit Cuevas M2 (Group 1)")
)
image(matrix(coefs[,1] + coefs[,3], nq, nq, byrow = TRUE),
  col = colors, breaks = breaks,
  axes = FALSE, main = paste("LPK Fit Cuevas M2 (Group 1)")
)
color_values <- cut(wy_pvals, breaks = 100, labels = FALSE)
colors <- viridis::viridis(100)
image(matrix(wy_pvals, nrow = nq, ncol = nq, byrow = TRUE),
  col = colors,
  axes = FALSE,
  main = "Westfall-Young Adjusted P-values"
)

```

<!-- }}} WY adjusted p-values  -->
